Toward Resolved Simulations of Burning Fronts in Thermonuclear X-ray Bursts

### 2018-09-29 astronum_2018/astronum-xrb.tex
some minor edits

The 2D calculations shown in the previous section require about 10k
node-hours of computational time to get to about 10~ms (those
particular simulations were run at NERSC).  We expect qualitatively different
results in 3D calculations: for example, the shear from the hurricane structure
acting along the flame front may cause instabilities that can affect the flame
propagation.
With our current simulation methodology,
a 3D version of this calculation is estimated to require 5M node-hours
on CPUs---this is about the size of an annual \MarginPar{double check
  \#s} allocation.
%Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
We plan to test this soon.
This simulation covers less than
1\% of the neutron star surface, so a fully resolved calculation of
burning over the entire neutron star is out of reach: a naive scaling
calculation suggest that 15 years of Moore's Law improvements are required.
%
% The calculation above is
% log_2 [ (4 pi 10^2) / (1.23^2) ] * 1.5
% for a NS of 10km radius vs the computational domain of 1.23km^2
%

Improvements are needed to push the scale of simulations
to the point where we can model XRBs at the full star scale with
resolved nuclear physics.
As the computational cost is roughly a product of the number of zones, the
number of zone updates, and the cost per zone update, we need to look for
improvements in each of these. The number of zones depends on the physical
model employed, the meshing strategy, and the accuracy required. The number of
zone updates depends on the smallest zone size and the largest characteristic
speed in the model. The computational cost per zone update is dominated by the
time spent in the reactions\MarginPar{verify fraction}. We discuss a few
possible improvements that can come both from implementing new algorithms and
from exploiting new hardware.

>>>>

The 2D calculations shown in the previous section require about 10k
node-hours of computational time to get to about 10~ms (running
on the Edison machine at NERSC).  We expect qualitatively different
results in 3D calculations: for example, the shear from the hurricane structure
acting along the flame front may cause instabilities that can affect the flame
propagation.
With our current simulation methodology,
a 3D version of this calculation is estimated to require 10M node-hours
or more---this is about the size of an annualallocation.
%Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
This simulation covers less than
1\% of the neutron star surface, so a fully resolved calculation of
burning over the entire neutron star is out of reach: a naive scaling
calculation suggest that 15 years of Moore's Law improvements are required.
%
% The calculation above is
% log_2 [ (4 pi 10^2) / (1.23^2) ] * 1.5
% for a NS of 10km radius vs the computational domain of 1.23km^2
%

Improvements are needed to push the scale of simulations
to the point where we can model XRBs at the full star scale with
resolved nuclear physics.
As the computational cost is roughly a product of the number of zones, the
number of zone updates, and the cost per zone update, we need to look for
improvements in each of these. The number of zones depends on the physical
model employed, the meshing strategy, and the accuracy required. The number of
zone updates depends on the smallest zone size and the largest characteristic
speed in the model. The computational cost per zone update is dominated by the
time spent in the reactions and equation of state. We discuss a few
possible improvements that can come both from implementing new algorithms and
from exploiting new hardware.
